### 2020-02-04 ICML/miaoyan/ordinalT_supp_v2.tex
Add files via upload

\documentclass[11pt]{article}

\usepackage{fancybox}

\usepackage{color}
\usepackage{url}
\usepackage[margin=1in]{geometry}

\renewcommand{\textfraction}{0.0}
\renewcommand{\topfraction}{1.0}
%\renewcommand{\textfloatsep}{5mm}

\usepackage{comment}
% Definitions of handy macros can go here
\usepackage{amsmath,amssymb,amsthm,bm,mathtools}
\usepackage{multirow}
\usepackage{natbib}
%\usepackage{dsfont,multirow,hyperref,setspace,natbib,enumerate}
\usepackage{dsfont,multirow,hyperref,setspace,enumerate}
\hypersetup{colorlinks,linkcolor={blue},citecolor={blue},urlcolor={red}}
\usepackage{algpseudocode,algorithm}
\algnewcommand\algorithmicinput{\textbf{Input:}}
\algnewcommand\algorithmicoutput{\textbf{Output:}}
\algnewcommand\INPUT{\item[\algorithmicinput]}
\algnewcommand\OUTPUT{\item[\algorithmicoutput]}

\mathtoolsset{showonlyrefs=true}

\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{pro}{Property}
\newtheorem{assumption}{Assumption}

\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{cor}{Corollary}
\newtheorem{example}{Example}
\newtheorem{rmk}{Remark}

\renewcommand{\thefigure}{{S\arabic{figure}}}%
\renewcommand{\thetable}{{S\arabic{table}}}%
\renewcommand{\figurename}{{Supplementary Figure}}
\renewcommand{\tablename}{{Supplementary Table}}
\setcounter{figure}{0}
\setcounter{table}{0}

\def\MLET{\hat \Theta_{\text{MLE}}}
\newcommand{\cmt}[1]{{\leavevmode\color{red}{#1}}}

\usepackage{dsfont}

\usepackage{multirow}

\DeclareMathOperator*{\minimize}{minimize}

\usepackage{wrapfig}

\usepackage{mathtools}
\mathtoolsset{showonlyrefs}
\newcommand*{\KeepStyleUnderBrace}[1]{%f
  \mathop{%
    \mathchoice
    {\underbrace{\displaystyle#1}}%
    {\underbrace{\textstyle#1}}%
    {\underbrace{\scriptstyle#1}}%
    {\underbrace{\scriptscriptstyle#1}}%
  }\limits
}
\usepackage{xr}
\externaldocument{ordinalT}
\input macros.tex

\title{Supplements for ``Tensor denoising and completion based on ordinal observations''}

%\author{%
%Chanwoo Lee \\
%University of Wisconsin -- Madison\\
 %\texttt{chanwoo.lee@wisc.edu} \\
%\And
%Miaoyan Wang \\
%University of Wisconsin -- Madison\\
%\texttt{miaoyan.wang@wisc.edu} \\
%}

\begin{document}

\begin{center}
\begin{spacing}{1.5}
\textbf{\Large Supplements for ``Tensor denoising and completion based on ordinal observations''}
\end{spacing}
\end{center}

\section{Proofs}
\subsection{Estimation error for tensor denoising}
\begin{proof}[Proof of Theorem~\ref{thm:rate}]
We suppress the subscript $\Omega$ in the proof, because the tensor denoising assumes complete observation $\Omega=[d_1]\times \cdots \times [d_K]$. It follows from the expression of $\flogl(\Theta)$ that
\begin{align}\label{eq:property}
{\partial \flogl\over \partial \theta_\omega}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_{\omega}=\ell\}}
{\dot{g}_\ell(\theta_\omega)\over g_\ell(\theta_\omega)},\notag\\
{\partial^2 \flogl\over \partial \theta_\omega^2}&=\sum_{\ell\in[L]}\mathds{1}_{\{y_\omega=\ell\}}{\ddot{g}_\ell(\theta_\omega)g_\ell(\theta_\omega)-\dot{g}^2_\ell(\theta_\omega)\over g^2_\ell(\theta_\omega)}\ \text{and}\quad
{\partial^2 \flogl\over \partial \theta_\omega \theta_\omega'}=0\ \text{if}\ \omega\neq \omega',
\end{align}
for all $\omega\in[d_1]\times \cdots \times [d_K]$.
Define $d_{\text{total}}=\prod_k d_k$. Let $\fplogl\in\mathbb{R}^{d_1\times\cdots\times d_K}$ denote the tensor of gradient with respect to $\Theta\in\mathbb{R}^{d_1\times \cdots\times d_K}$, and $\fpplogl$ the corresponding Hession matrix of size $d_\text{total}$-by-$d_{\text{total}}$. Here, $\Vec(\cdot)$ denotes the operation that turns a tensor into a vector. By~\eqref{eq:property}, $\fpplogl$ is a diagonal matrix. Recall that
\[
U_\alpha=\max_{\ell\in[L],|\theta|\leq \alpha}{\dot{g}_\ell(\theta)\over g_\ell(\theta)}>0 \quad \text{and}\quad
L_\alpha=\min_{\ell\in[L],|\theta|\leq \alpha} {\dot{g}^2_\ell(\theta)-\ddot{g}_\ell(\theta)g_\ell(\theta)\over g^2_\ell(\theta)}>0.
\]
Therefore, all entries in $\fplogl$ are upper bounded $U_\alpha>0$, and all diagonal entries in $\fpplogl$ are upper bounded by $-L_{\alpha}<0$.

By the second-order Taylor's expansion of $\flogl(\Theta)$ around $\trueT$, we obtain
\begin{equation}\label{eq:taylor}
\flogl(\Theta)=\flogl(\trueT)+\langle\Vec(\fplogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT),
\end{equation}
$\check\Theta=\gamma\trueT+(1-\gamma)\Theta$ for some $\gamma\in[0,1]$, and $\fpplogl(\check\Theta)$ denotes the $\prod_kd_k$-by-$\prod_k d_k$ Hession matrix evaluated at $\check\Theta$.

We first bound the linear term in~\eqref{eq:taylor}. Note that, by Lemma~\ref{lem:inq},
\begin{equation}\label{eq:linear}
|\langle\Vec(\fplogl(\trueT), \Vec(\Theta-\trueT)  \rangle|\leq \snormSize{}{\fplogl(\trueT)} \nnormSize{}{\Theta-\trueT},
\end{equation}
where $\snormSize{}{\cdot}$ denotes the tensor spectral norm and $\nnormSize{}{\cdot}$ denotes the tensor nuclear norm. Define
\[
s_\omega={\partial \tL_\tY\over \partial \theta_\omega}\Big|_{\Theta=\trueT} \;\; \textrm{ for all } \; \omega\in[d_1]\times\cdots\times [d_K].
\]
Based on~\eqref{eq:property} and the definition of $U_\alpha$, $\fplogl(\trueT)=\entry{s_{\omega}}$ is a random tensor whose entries are independently distributed satisfying
\begin{equation}\label{eq:norm}
\mathbb{E}(s_\omega)=0,\quad |s_\omega|\leq U_\alpha, \quad \text{for all }\omega\in[d_1]\times \cdots \times [d_K].
\end{equation}
By lemma~\ref{lem:noisytensor}, with probability at least $1-\exp(-C_1 \sum_kd_k)$, we have
\begin{equation}\label{eq:normrandom}
\snormSize{}{\fplogl(\trueT)} \leq C_2 U_\alpha\sqrt{\sum_k d_k},
\end{equation}
where $C_1, C_2$ are two positive constants that depend only on $K$. Furthermore, note that $\text{rank}(\Theta)\leq \mr$, $\text{rank}(\trueT)\leq \mr$, so $\text{rank}(\Theta-\trueT)\leq 2\mr$. By lemma~\ref{lem:nuclear}, $\nnormSize{}{\Theta-\trueT}\leq (2r_{\max})^{K-1\over 2}\FnormSize{}{\Theta-\trueT}$. Combining~\eqref{eq:linear}, \eqref{eq:norm} and \eqref{eq:normrandom}, we have that, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
\begin{equation}\label{eq:linearconclusion}
|\langle \Vec(\fplogl(\trueT)), \Vec(\Theta-\trueT)  \rangle | \leq C_2 U_\alpha  \sqrt{r_{\max}^{K-1} \sum_k d_k}  \FnormSize{}{\Theta-\trueT}.
\end{equation}

We next bound the quadratic term in \eqref{eq:taylor}. Note that
\begin{align}\label{eq:quadratic}
 \Vec(\Theta-\trueT)^T \fpplogl(\check{\Theta})\Vec(\Theta-\trueT)&=\sum_\omega \left( {\partial^2\tL_{\tY}\over \partial \theta^2_\omega} \Big|_{\Theta=\check\Theta} \right)(\theta_\omega-\theta_{{\text{true}},\omega})^2 \nonumber \\
&\leq - L_\alpha\sum_{\omega}(\Theta_{\omega}-\Theta_{\text{true},\omega})^2 \nonumber \\
&=-L_\alpha\FnormSize{}{\Theta-\trueT}^2,
\end{align}
where the second line comes from the fact that  $\mnormSize{}{\check\Theta}\leq \alpha$ and the definition of $L_\alpha$.

Combining~\eqref{eq:taylor}, \eqref{eq:linearconclusion} and~\eqref{eq:quadratic}, we have that, for all $\Theta\in\tP$, with probability at least $1-\exp(-C_1 \sum_kd_k)$,
\[
\tL_\tY(\Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha  \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\Theta-\trueT}^2.
\]
In particular, the above inequality also holds for $\hat \Theta\in\tP$. Therefore,
\[
\tL_\tY(\hat \Theta)\leq \tL_{\tY}(\trueT)+C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2} \FnormSize{}{\hat \Theta-\trueT}^2.
\]
Since $\hat \Theta=\arg\max_{\Theta\in\tP}\tL_\tY(\Theta)$, $\tL_\tY(\hat \Theta)-\tL_{\tY}(\trueT)\geq 0$, which gives
\[
C_2U_\alpha \left(r_{\max}^{K-1}\sum_k d_k\right)^{1/2}  \FnormSize{}{\hat \Theta-\trueT}-{L_\alpha\over 2}\FnormSize{}{\hat \Theta-\trueT}^2\geq 0.
\]
Henceforth,
\[
{1\over \sqrt{\prod_k d_k}} \FnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\over L_\alpha \sqrt{\prod_k d_k}}={2C_2U_\alpha r_{\max}^{(K-1)/2}\over L_\alpha} \sqrt{ \sum_k d_k \over \prod_k d_k}.
\]
This completes the proof.
\end{proof}

\begin{proof}[Proof of Corollary~\ref{cor:prediction}]
The result follows immediately from Theorem~\ref{thm:rate} and Lemma~\ref{lem:KL}.
\end{proof}

\subsection{Sample complexity for tensor completion}
\begin{proof}[Proof of Theorem~\ref{thm:minimax}]

Let $d_{\text{total}}=\prod_{k\in[K]}d_k$, and $\gamma\in[0,1]$ be a constant to be specified later.  Our strategy is to construct a finite set of tensors $\tX=\{\Theta_i\colon i=1,\ldots \}\subset \tP$ satisfying the properties of (i)-(iv) in Lemma~\ref{lem:construction}. By Lemma~\ref{lem:construction}, such a subset of tensors exist. For any tensor  $\Theta\in\tX$, let $\mathbb{P}_{\Theta}$ denote the distribution of $\tY|\Theta$, where $\tY$ is the ordinal tensor. In particular, $\mathbb{P}_{\mathbf{0}}$ is the distribution of $\tY$ induced by the zero parameter tensor $\mathbf{0}$, i.e., the distribution of $\tY$ conditional on the parameter tensor $\Theta=\mathbf{0}$. Based on the Remark for Lemma~\ref{lem:KL}, we have
\begin{equation}\label{eq:KLbound1}
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\mathbf{0}})\leq C \FnormSize{}{\Theta}^2,
\end{equation}
where $C={(4L-6) \dot{f}^2(0)\over  A_\alpha}>0$ is a constant independent of the tensor dimension and rank.
Combining the inequality~\eqref{eq:KLbound1} with property (iii) of $\tX$, we have
\begin{equation}\label{eq:KLbound}
\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\mathbf{0}})\leq \gamma^2 R_{\max} d_{\max}.
\end{equation}
From~\eqref{eq:KLbound} and the property (i), we deduce that the condition
\begin{equation}\label{eq:totalKL}
{1\over \text{Card}(\tX)-1}\sum_{\Theta \in\tX}\text{KL}(\mathbb{P}_{\Theta}, \mathbb{P}_{\mathbf{0}})\leq \varepsilon \log_2\left\{\text{Card}(\tX)-1 \right\}
\end{equation}
holds for any $ \varepsilon \geq 0$ when $\gamma\in[0,1]$ is chosen to be sufficiently small depending on $\varepsilon$, e.g., $\gamma \leq \sqrt{3\varepsilon}$. By applying Lemma~\ref{lem:Tsybakov} to~\eqref{eq:totalKL}, and in view of the property (iv), we obtain that
\begin{equation}\label{eq:final}
\inf_{\hat \Theta}\sup_{\trueT\in \tX}\mathbb{P}\left(\FnormSize{}{\hat \Theta- \trueT}\geq  {\gamma\over 8} \min\left\{ \alpha\sqrt{d_{\text{total}}}, C^{-1/2}\sqrt{ R_{\max}d_{\max}}\right\} \right)\geq {1\over 2}\left(1-2\varepsilon-\sqrt{16 \varepsilon \over R_{\max}d_{\max}}\right).
\end{equation}
Note that $\text{Loss}(\hat \Theta, \trueT)=\FnormSize{}{\hat \Theta- \trueT}^2/d_{\text{total}}$ and $\tX\subset \tP$. By taking $\varepsilon=1/8$ and $\gamma=1/2$, we conclude from~\eqref{eq:final} that
\begin{equation*}\label{eq:prob}
\inf_{\hat \Theta}\sup_{\trueT\in \tP}\mathbb{P}\left(\text{Loss}(\hat \Theta, \trueT)\geq {1 \over 256}\min\left \{ \alpha^2,  {C^{-1}R_{\max}d_{\max}\over d_{\text{total}}}\right \}\right)\geq {1\over 2}\left({3\over 4}- {2\over R_{\max}d_{\max}} \right)\geq {1\over 8}.
\end{equation*}
This completes the proof.
\end{proof}

\begin{proof}[Proof of Theorem~\ref{thm:completion}]

For notational convenience, we use $\MFnormSize{}{\Theta}=\sum_{\omega\in\Omega}\Theta^2_\omega$ to denote the sum of squared entries over the observed set $\Omega$, for a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$.

Following a similar argument as in the proof of Theorem~\ref{thm:rate}, we have
\begin{equation}\label{eq:Taylor2}
\logl(\Theta)=\logl(\trueT)+\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle+{1\over 2}\Vec(\Theta-\trueT)^T\pplogl(\check\Theta)\Vec(\Theta-\trueT),
\end{equation}
where
\begin{enumerate}
\item $\plogl$ is a $d_1\times\cdots\times d_K$ tensor with $|\Omega|$ nonzero entries, and each entry is upper bounded by $U_\alpha>0$.
\item $\pplogl$ is a diagonal matrix of size $d_{\text{total}}$-by-$d_{\text{total}}$ with $|\Omega|$ nonzero entries, and each entry is upper bounded by $-L_{\alpha}<0$.
\end{enumerate}

Similar to~\eqref{eq:linear} and~\eqref{eq:quadratic}, we have
\begin{equation}\label{eq:linear2}
|\langle\Vec(\plogl),\ \Vec(\Theta-\trueT)\rangle|\leq C_2U_\alpha \sqrt{r_{\max}^{K-1}\sum_k d_k}\MFnormSize{}{\Theta-\trueT}
\end{equation}
and
\begin{equation}\label{eq:quadratic2}
\Vec(\Theta-\trueT)^T\fpplogl(\check\Theta)\Vec(\Theta-\trueT)\leq -L_\alpha \MFnormSize{}{\Theta-\trueT}^2.
\end{equation}

Combining~\eqref{eq:Taylor2}-\eqref{eq:quadratic2} with the fact that $\logl(\hat \Theta)\geq \logl(\trueT)$, we have
\begin{equation}\label{eq:sample}
\MFnormSize{}{\hat \Theta-\trueT}\leq {2C_2U_\alpha  r_{\max}^{(K-1)/2} \over L_\alpha} \sqrt{\sum_kd_k}.
\end{equation}
Lastly, we invoke the result regarding the closeness of $\Theta$ to its sampled version $\Theta_{\Omega}$, under the entrywise bound condition. Note that $\mnormSize{}{\hat\Theta-\trueT}\leq 2\alpha$ and $\text{rank}(\hat \Theta-\trueT)\leq 2\mr$. By Lemma~\ref{lem:Mnormbound}, $\anormSize{}{\hat \Theta-\trueT}\leq 2^{(3K-1)/2} \alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$. Therefore, the condition in Lemma~\ref{lem:convexity} holds with $\beta=2^{(3K-1)/2}\alpha \left({\prod r_k \over r_{\max}}\right)^{3/2}$. Applying Lemma~\ref{lem:convexity} to~\eqref{eq:sample} gives
\begin{align}
 \PiFnormSize{}{\hat \Theta-\trueT}^2&\leq {1\over m}\MFnormSize{}{\hat \Theta-\trueT}^2+c\beta\sqrt{\sum_k d_k\over |\Omega|}\\
 &\leq {C_2  r^{K-1}_{\max}} {\sum_k d_k \over |\Omega|}+C_1 \alpha r_{\max}^{3(K-1)/2}\sqrt{\sum_kd_k\over |\Omega|},
\end{align}
with probability at least $1-\exp(-{\sum_kd_k\over \sum_k \log d_k})$ over the sampled set $\Omega$. Here $C_1, C_2>0$ are two constants independent of the tensor dimension and rank. Therefore,
\[
 \PiFnormSize{}{\hat \Theta-\trueT}^2\to 0,\quad \text{as}\quad {|\Omega|\over \sum_kd_k}\to \infty,
\]
provided that $r_{\max}=O(1)$.
\end{proof}

\subsection{Auxiliary lemmas}
\label{sec:lemma}

We begin with various notion of tensor norms that are useful for the proofs of the main theorems.

\begin{defn}[Atomic M-norm~\citep{ghadermarzy2019near}]
Define $T_{\pm}=\{\tT\in\{\pm 1 \}^{d_1\times \cdots \times d_K}\colon \text{rank}(\tT)=1\}$. The atomic M-norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
\begin{align}
\anormSize{}{\Theta}&=\inf\{t>0: \Theta\in t\text{conv}(T_{\pm})\}\\
&=\inf\left\{\sum_{\tX\in T_{\pm}}c_x\colon \Theta=\sum_{\tX\in T_{\pm}} c_x\tX, \ c_x>0\right\}.
\end{align}
\end{defn}

\begin{defn}[Spectral norm~\citep{lim2005singular}]
The spectral norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
\[
\snormSize{}{\Theta}=\sup\left\{\langle \Theta, \mx_1\otimes \cdots \otimes \mx_K\rangle \colon \vnormSize{}{\mx_k}=1,\ \mx_k\in\mathbb{R}^{d_k},\ \text{for all}\ k\in[K]\right\}.
\]
\end{defn}

\begin{defn}[Nuclear norm~\citep{friedland2018nuclear}]
The nuclear norm of a tensor $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is defined as
\[
\nnormSize{}{\Theta}=\inf
\left\{
\sum_{i\in[r]}|\lambda_i|\colon \Theta=\sum_{i=1}^r \lambda_i\mx^{(i)}_1\otimes \cdots \otimes \mx^{(i)}_K,\ \vnormSize{}{\mx^{(i)}_k}=1,\ \mx^{(i)}_k\in\mathbb{R}^{d_k},\ \text{for all} k\in[K],\ i\in[r]
\right\},
\]
where the infimum is taken over all $r\in\mathbb{N}$ and $\norm{\mx^{(i)}_k}=1$ for all $i\in[r]$ and $k\in[K]$.
\end{defn}

\begin{lem}[M-norm and infinity norm~\citep{ghadermarzy2019near}]\label{lem:Mnormbound}
Let $\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}$ be an order-$K$, rank-$(r_1,\ldots,r_K)$ tensor. Then
\[
\mnormSize{}{\Theta}\leq \anormSize{}{\Theta}\leq \left(\prod_k r_k \over r_{\max}\right)^{3\over 2} \mnormSize{}{\Theta}.
\]
\end{lem}

\begin{lem}[Nuclear norm and F-norm] \label{lem:nuclear}
Let $\tA\in\mathbb{R}^{d_1\times\cdots\times d_K}$ be an order-$K$ tensor with Tucker $\text{rank}(\tA)=(r_1,\ldots,r_K)$. Then
\[
\nnormSize{}{\tA} \leq \sqrt{\prod_k r_k\over \max_k r_k} \FnormSize{}{\tA},
\]
where $\nnormSize{}{\cdot}$ denotes the nuclear norm of the tensor.
\end{lem}

\begin{proof}
Without loss of generality, suppose $r_1=\min_k r_k$. Let $\tA_{(k)}$ denote the mode-$k$ matricization of $\tA$ for all $k\in[K]$. By \citet[Corollary 4.11]{wang2017operator}, and the invariance relationship between a tensor and its Tucker core~\citep[Section 6]{jiang2017tensor}, we have
\begin{equation}\label{eq:norminequality}
\nnormSize{}{\tA} \leq \sqrt{\prod_{k\geq 2} r_k \over \max_{k\geq 2} r_k} \nnormSize{}{\tA_{(1)}},
\end{equation}
where $\tA_{(1)}$ is a $d_1$-by-$\prod_{k\geq 2}d_k$ matrix with matrix rank $r_1$. Furthermore, the relationship between the matrix norms implies that $\nnormSize{}{\tA_{(1)}}\leq \sqrt{r_1}\FnormSize{}{\tA_{(1)}}=\sqrt{r_1}\FnormSize{}{\tA}$. Combining this fact with the inequality~\eqref{eq:norminequality} yields the final claim.
\end{proof}

\begin{lem} \label{lem:inq}
Let $\tA, \; \tB$ be two order-$K$ tensors of the same dimension. Then
\[
|\langle \tA,\tB\rangle| \leq \snormSize{}{\tA}   \nnormSize{}{\tB}.
\]
\end{lem}

\begin{proof}
By~\citet[Proposition 3.1]{friedland2018nuclear}, there exists a nuclear norm decomposition of $\tB$, such that
\[
\tB=\sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r,\quad \ma_r^{(k)}\in\mathbf{S}^{d_k-1}(\mathbb{R}),\quad \text{for all }k\in[K],
\]
and $\nnormSize{}{\tB}=\sum_{r}|\lambda_r|$. Henceforth we have
\begin{align*}
|\langle \tA,\tB\rangle|&=| \langle \tA, \sum_{r} \lambda_r \ma^{(1)}_r\otimes \cdots\otimes \ma^{(K)}_r \rangle|\leq \sum_r |\lambda_r| |\langle \tA, \ma^{(1)}_r \otimes \cdots\otimes \ma^{(K)}_r \rangle|\\
&\leq \sum_{r}|\lambda_r| \snormSize{}{\tA}= \snormSize{}{\tA}\nnormSize{}{\tB},
\end{align*}
which completes the proof.
\end{proof}

The following lemma provides the bound on the spectral norm of random tensors. The result was firstly presented in~\cite{nguyen2015tensor}, and we adopt the version from~\cite{tomioka2014spectral}.
\begin{lem}[\cite{tomioka2014spectral}]\label{lem:tensor}
Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
\[
\mathbb{E}(s_{\omega})=0,\quad \text{and} \quad\mathbb{E}(e^{ts_{\omega}})\leq e^{t^2L^2/2}.
\]
Then the spectral norm $\snormSize{}{\tS}$ satisfies that,
\[
\snormSize{}{\tS}\leq \sqrt{{8L^2} \log (12K) \sum_k d_k +\log (2/\delta)},
\]
with probability at least $1-\delta$.
\end{lem}

\begin{lem} \label{lem:noisytensor}
Suppose that $\tS=\entry{s_{\omega}}\in\mathbb{R}^{d_1\times \cdots \times d_K}$ is an order-$K$ tensor whose entries are independent random variables that satisfy
\[
\mathbb{E}(s_{\omega})=0,\quad \text{and}\quad |s_{\omega}|\leq U.
\]
Then we have
\[
\mathbb{P}\left(\snormSize{}{\tS}\geq C_2 U\sqrt{\sum_k d_k} \right)\leq \exp\left(-C_1  \log K \sum_k d_k\right)
\]
where $C_1>0$ is an absolute constant, and $C_2>0$ is a constant that depends only on $K$.
\end{lem}

\begin{proof}  Note that the random variable $U^{-1}s_{\omega}$ is zero-mean and supported on $[-1,1]$. Therefore, $U^{-1}s_{\omega}$ is sub-Gaussian with parameter ${1-(-1)\over 2}=1$, i.e.
\[
\mathbb{E}(U^{-1}s_{\omega})=0,\quad \text{and}\quad \mathbb{E}(e^{tU^{-1}s_{\omega}})\leq e^{t^2/2}.
\]
It follows from Lemma~\ref{lem:tensor} that, with probability at least $1-\delta$,
\[
\snormSize{}{U^{-1}\tS}\leq \sqrt{\left(c_0\log K+c_1\right) \sum_k d_k +\log (2/\delta)},
\]
where $c_0, c_1>0$ are two absolute constants. Taking $\delta=\exp (-C_1\log K \sum_k d_k)$ yields the final claim, where $C_2=c_0\log K+c_1+1>0$ is another constant.
\end{proof}

\begin{lem}\label{lem:KLentry} Let $X,\; Y$ be two discrete random variables taking values on $L$ possible categories, with category probabilities $\{p_\ell\}_{\ell\in[L]}$ and $\{q_\ell\}_{\ell\in[L]}$, respectively.  Suppose $p_\ell$, $q_\ell>0$ for all $i\in[L]$. Then, the Kullback-Leibler (KL) divergence satisfies that
\[
\text{KL}(X||Y)\stackrel{\text{def}}{=}-\sum_{\ell\in[L]}\mathbb{P}_X(\ell)\log\left\{{\mathbb{P}_Y(\ell)\over \mathbb{P}_X(\ell)}\right\} \leq \sum_{\ell \in [L]}{(p_\ell-q_\ell)^2 \over q_\ell}.
\]
\end{lem}
\begin{proof} Using the fact $\log x\leq x-1$ for $x>0$, we have that
\begin{align}\label{eq:KL}
\text{KL}(X||Y)&=\sum_{\ell\in[L]}p_\ell\log{p_\ell\over q_\ell}\\
&\leq \sum_{\ell\in[L]} {p_\ell\over q_\ell}(p_\ell-q_\ell)\\
&=\sum_{\ell\in[L]} \left({p_\ell\over q_\ell }- 1\right)(p_\ell-q_\ell)+ \sum_{\ell\in[L]} (p_\ell-q_\ell).
\end{align}
Note that $\sum_{\ell\in[L]}(p_\ell-q_\ell)=0$. Therefore,
\[
\text{KL}(X||Y)\leq \sum_{\ell\in[L]}\left( {p_\ell\over q_\ell}-1\right)\left(p_\ell-q_\ell\right)=\sum_{\ell\in[L]}{(p_\ell-q_\ell)^2\over q_\ell}.
\]
\end{proof}

\begin{lem}[KL divergence and F-norm]~\label{lem:KL}
Let $\tY\in[L]^{d_1\times \cdots \times d_K}$ be an ordinal tensor generated from the model~\eqref{eq:model} with the link function $f$ and parameter tensor $\Theta$. Let $\mathbb{P}_{\Theta}$ denote the joint categorical distribution of $\tY|\Theta$ induced by the parameter tensor $\Theta$, where $\mnormSize{}{\Theta}\leq \alpha$. Define
\begin{equation}\label{eq:ass}
A_\alpha=\min_{\ell\in[L], |\theta|\leq \alpha}\left[f(\theta+b_\ell)-f(\theta+b_{\ell-1})\right].
\end{equation}
Then, for any two tensors $\Theta,\; \Theta^*$ in the parameter spaces, we have
\[
KL(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*})\leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{lem}
\begin{proof} Suppose that the distribution over the ordinal tensor $\tY=\entry{y_\omega}$ is induced by $\Theta=\entry{\theta_\omega}$. Then, based on the generative model~\eqref{eq:model},
\[
\mathbb{P}(y_\omega=\ell | \theta_\omega)=f(\theta_\omega+b_{\ell})-f(\theta_\omega+b_{\ell-1}),
\]
for all $\ell\in[L]$ and $\omega\in[d_1]\times \cdots \times [d_K]$. For notational convenience, we suppress the subscribe in $\theta_\omega$ and simply write $\theta$ (and respectively, $\theta^*$). Based on Lemma~\ref{lem:KLentry} and Taylor expansion,
\begin{align}
\text{KL}(\theta|| \theta^*) & \leq \sum_{\ell\in[L]} {\left[f(\theta+b_{\ell})-f(\theta+b_{\ell-1})-f(\theta^*+b_{\ell})+f(\theta^*+b_{\ell-1})\right]^2\over f(\theta^*+b_{\ell})-f(\theta^*+b_{\ell-1})}\\
 &\leq \sum_{\ell=2}^{L-1} {\left[\dot{f}(\eta_\ell+b_\ell)-\dot{f}(\eta_{\ell-1}+b_{\ell-1})\right]^2 \over f(\theta^*+b_\ell)-f(\theta^*+b_{\ell-1})} \left(\theta-\theta^*\right)^2+{\dot{f}^2(\eta_1+b_1) \over f(\theta^*+b_1)} (\theta-\theta^*)^2\\
 &\quad \quad\quad \quad  +{\dot{f}^2(\eta_{L-1}+b_{L-1})\over 1-f(\theta^*+b_{L-1})}(\theta-\theta^*)^2,
\end{align}
where $\eta_\ell$ and $\eta_{\ell-1}$ fall between $\theta$ and $\theta^*$. Therefore,
\begin{equation}\label{eq:entrywise}
\text{KL}(\theta|| \theta^*) \leq \left({4(L-2)\over A_\alpha}+ {2\over A_\alpha}\right)\dot{f}^2(0)(\theta-\theta^*)^2={2(2L-3)\over A_\alpha} \dot{f}^2(0)(\theta-\theta^*)^2,
\end{equation}
where we have used Taylor expansion, the bound~\eqref{eq:ass}, and the fact that $\dot{f}(\cdot)$ peaks at zero for an unimodal and symmetric function. Now summing~\eqref{eq:entrywise} over the index set $\omega\in[d_1]\times \cdots \times [d_K]$ gives
\[
\text{KL}(\mathbb{P}_{\Theta}|| \mathbb{P}_{\Theta^*}) =\sum_{\omega\in[d_1]\times \cdots \times[d_K]} \text{KL}(\theta_\omega || \theta^*_\omega) \leq {2(2L-3)\over A_\alpha} \dot{f}^2(0)\FnormSize{}{\Theta-\Theta^*}^2.
\]
\end{proof}

\begin{rmk} In particular, let $\mathbb{P}_{\bf{0}}$ denote the distribution of $\tY|\bf{0}$ induced by the zero parameter tensor. Then we have
\[
\text{KL}(\mathbb{P}_{\Theta}||\mathbb{P}_{\bf{0}})\leq {2(2L-3)\over A_\alpha}  \dot{f}^2(0)\FnormSize{}{\Theta}^2.
\]
\end{rmk}

\begin{lem}\label{lem:construction}
Assume the same setup as in Theorem~\ref{thm:minimax}. Without loss of generality, suppose $d_1=\max_kd_k$. Define $R=\max_k r_k$ and $d_{\text{total}}=\prod_{k\in[K]} d_k$. For any constant $0\leq \gamma \leq 1$, there exist a finite set of tensors $\tX=\{\Theta_i: i=1,\ldots\}\subset \tP$ satisfying the following four properties:
\begin{enumerate}[(i)]
\item $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$, where $\text{Card}$ denotes the cardinality;
\item $\tX$ contains the zero tensor $\mathbf{0}\in\mathbb{R}^{d_1\times \cdots\times d_K}$;
\item $\mnormSize{}{\Theta}\leq \gamma \min\left\{ \alpha ,\ C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\} $ for any element $\Theta\in\tX$;
\item $\FnormSize{}{\Theta_i-\Theta_j}\geq {\gamma\over 4} \min\left\{ \alpha\sqrt{d_{\text{total}}},\ C^{-1/2}\sqrt{Rd_1}\right\}$ for any two distinct elements $\Theta_i,\; \Theta_j\in\tX$,
\end{enumerate}
Here $C=C(\alpha,L,f,\mb)={(4L-6)\dot{f}^2(0)\over A_\alpha }>0$ is a constant independent of the tensor dimension and rank.
\end{lem}

\begin{proof}
Given a constant $0\leq \gamma \leq 1$, we define a set of matrices:
\[
\tC=\left\{\mM=(m_{ij})\in\mathbb{R}^{d_1\times R}: a_{ij}\in \left\{ 0,\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}} \right\}\right\} ,\  \forall (i,j)\in[d_1]\times[R]\right\}.
\]
We then consider the associated set of block tensors:
\begin{align}
\tB=\tB(\tC)=\{\Theta\in\mathbb{R}^{d_1\times \cdots \times d_K}\colon& \Theta=\mA\otimes \mathbf{1}_{d_3}\otimes \cdots \otimes \mathbf{1}_{d_K}, \\
&\ \text{where}\ \mA=(\mM|\cdots|\mM|\mO) \in\mathbb{R}^{d_1\times d_2},\ \mM\in\tC\},
\end{align}
where $\mathbf{1}_d$ denotes a length-$d$ vector with all entries 1, $\mO$ denotes the $d_1\times (d_2-R\lfloor d_2/R \rfloor)$ zero matrix, and $\lfloor d_2/ R \rfloor$ is the integer part of $d_2/R$. In other words, the subtensor $\Theta(\mI, \mI,i_3, \ldots,i_K)\in\mathbb{R}^{d_1\times d_2}$ are the same for all fixed $(i_3,\ldots,i_K)\in[d_3]\times \cdots \times [d_K]$, and furthermore, each subtensor $\Theta(\mI,\mI, i_3,\ldots,i_K)$ itself is filled by copying the matrix $\mM\in\mathbb{R}^{d_1\times R}$ as many times as would fit.

By construction, any element of $\tB$, as well as the difference of any two elements of $\tB$, has Tucker rank at most $\max_k r_k\leq R$, and the entries of any tensor in $\tB$ take values in $[0,\alpha]$. Thus, $\tB\subset\tP$. By Lemma~\ref{lem:VGbound}, there exists a subset $\tX\subset \tB$ with cardinality $\text{Card}(\tX)\geq 2^{Rd_1/8}+1$ containing the zero $d_1\times \cdots \times d_K$ tensor, such that, for any two distinct elements $\Theta_i$ and $\Theta_j$ in $\tX$,
\[
\FnormSize{}{\Theta_i-\Theta_j}^2 \geq {Rd_1\over 8} \gamma^2\min\left\{ \alpha^2, {C^{-1}Rd_1 \over d_{\text{total}}}\right\} \lfloor {d_2\over R} \rfloor \prod_{k\geq 3}d_k\geq {\gamma^2\min\left\{ \alpha^2 d_{\text{total}}, C^{-1}Rd_1\right\}  \over 16}.
\]
In addition, each entry of $\Theta\in\tX$ is bounded by $\gamma \min\left\{ \alpha , C^{-1/2}\sqrt{Rd_1\over d_{\text{total}}}\right\} $. Therefore the Properties (i) to (iv) are satisfied.
\end{proof}

\begin{lem}[Varshamov-Gilbert bound]\label{lem:VGbound}
Let $\Omega=\{(w_1,\ldots,w_m)\colon w_i\in\{0,1\}\}$. Suppose $m>8$. Then there exists a subset $\{w^{(0)},\ldots,w^{(M)}\}$ of $\Omega$ such that $w^{(0)}=(0,\ldots,0)$ and
\[
\zeronormSize{}{w^{(j)}-w^{(k)}}\geq {m\over 8},\quad \text{for} \ 0\leq j<k\leq M,
\]
where $\zeronormSize{}{\cdot}$ denotes the Hamming distance, and $M\geq 2^{m/8}$.
\end{lem}

\begin{lem}[Theorem 2.5 in~\cite{tsybakov2009introduction}]\label{lem:Tsybakov}
Assume that a set $\tX$ contains element $\Theta_0, \Theta_1, \ldots,\Theta_M$ ($M\geq 2$) such that
\begin{itemize}
\item $d(\Theta_j,\ \Theta_k)\geq 2s>0$, $\forall 0\leq j\leq k\leq M$;
\item $\mathbb{P}_j\ll\mathbb{P}_0$, $\forall j=1,\ldots,M$, and
\[
{1\over M}\sum_{j=1}^M KL(\mathbb{P}_j||\mathbb{P}_0)\leq \alpha \log M
\]
where $d\colon \tX\times \tX\mapsto [0,+\infty]$ is a semi-distance function, $0<\alpha<{1/8}$ and $P_j=P_{\Theta_j}$, $j=0,1\ldots,M$.
\end{itemize}
Then
\[
\inf_{\hat \Theta}\sup_{\Theta\in\tX} \mathbb{P}_{\Theta}(d(\hat \Theta, \Theta)\geq s)\geq {\sqrt{M}\over 1+\sqrt{M}}\left(1-2\alpha-\sqrt{2\alpha\over \log M} \right)>0.
\]
\end{lem}

\begin{lem}[Lemma 28 in~\cite{ghadermarzy2019near}]\label{lem:convexity}
Define $\mathbb{B}_{M}(\beta)=\{\Theta\in \mathbb{R}^{d_1\times \cdots \times d_K}\colon \anormSize{}{\Theta}\leq \beta \}$.  Let $\Omega\subset[d_1]\times\cdots \times [d_K]$ be a random set with $m=|\Omega|$, and assume that each entry in $\Omega$ is drawn with replacement from $[d_1]\times\cdots\times[d_K]$ using probability $\Pi$. Define
\[
\PiFnormSize{}{\Theta}^2={1\over m}\mathbb{E}_{\Omega\in\Pi}\MFnormSize{}{\Theta}^2.
\]
Then, there exists a universal constant $c>0$, such that, with probability at least $1-\exp\left(-{\sum_kd_k \over \sum_k \log d_k} \right)$ over the sampled set $\Omega$,
\[
{1\over m}\MFnormSize{}{\Theta}^2 \geq \PiFnormSize{}{\Theta}^2-c\beta\sqrt{\sum_k d_k\over m}
\]
holds uniformly for all $\Theta\in\mathbb{B}_M(\beta)$.
\end{lem}

\section{Convexity of the likelihood function}
\begin{thm}\label{thm:convexity}
\begin{align}
 \logl(\Theta, \mb)&=\sum_{\omega\in\Omega}\sum_{\ell\in[L]} \Big\{\mathds{1}_{\{y_\omega=\ell\}} \log \big[f(\theta_\omega+b_\ell)-  f(\theta_\omega+b_{\ell-1})\big]\Big\}, \text{ where } f(x) = \frac{e^x}{1+e^x}.
 \end{align}
is a concave function to $(\Theta,\mb)$
\end{thm}
\begin{proof}
We use $(\Theta,\mb)$ abusively as $(Vec(\Theta)^T,\mb^T)^T$ in this proof.
Let us denote $\lambda(u,v) = \log \big[f(u)-  f(v)\big]$.
It is enough to show that $\lambda(u,v)$
is a concave function to $(u,v)$ where $u>v$. Because if $\lambda(u,v)$ is a concave function and $u,v$ are both linear functions of $(\Theta,\mb) $ such that
$$u = \ma_1^T(\Theta,\mb), v = \ma_2^T(\Theta,\mb) \text{ for some }\ma_1,\ma_2\in \mathbb{R}^{d_{\text{total}}+L-1},$$ then $\lambda(u,v) = \lambda(\ma_1^T(\Theta,\mb),\ma_2^T(\Theta,\mb)))$ is a concave function to $(\Theta,\mb)$
by the definition of the convexity. Furthermore, we can conclude that $\logl(\Theta, \mb) $ is a concave function to $(\Theta,\mb)$ because $\logl(\Theta, \mb)$
can be written as the form of summations of $\lambda(u,v)$. To prove the convexity of $\lambda(u,v)$, we have the following.
\begin{equation*}
\lambda(u,v) = \log\big[f(u)-f(v)\big]=\log\big[\int\mathds{1}_{(u,v)}(x)f'(x)dx\big].
\end{equation*}
Notice that  $\mathds{1}_{(u,v)}(x)$ is log-concave to $(u,v,x)$ and $f'(x)$ is log-concave to $x$ because $f'(x) = \frac{e^x}{(1+e^x)^2}$. Those convexities imply that $\mathds{1}_{(u,v)}(x)f'(x)$
is a log-concave function to $(u,v,x)$. Therefore, we can conclude that the $\lambda(u,v)$ is a concave function from the lemma~\ref{lem:convexity}.
\end{proof}

\begin{lem}[Corollary 3.5 in~\cite{brascamp2002extensions}]\label{lem:convexity}
Let $F(x,y):\mathbb{R}^{m+n}\rightarrow \mathbb{R}$ be an integrable function where $x\in \mathbb{R}^{m},y\in \mathbb{R}^n$. Let $$G(x) = \int_{\mathbb{R}^n}F(x,y)dy.$$ If $F(x,y)$ is a log concave function to (x,y), then $G(x)$ is a log concave function.
\end{lem}

\section{Brain clustering method and result}
\subsection{Clustering method}
We perform clustering based on the estimated latent parameter $\hat\Theta$. Our clustering method is based on multidimensional version of Principal Component Analysis (PCA) and Singular Value Decomposition (SVD). In matrices case, we perform clustering on a $m\times n$ data matrix $X$ based on the following procedure. First, we factorize $X$ into
\begin{equation*}
X = U\Sigma V^T,
\end{equation*}
where $\Sigma$ is a diagonal matrix and $ U,V$ are factor matrices with orthogonal columns. We interpret each column of $V$ as a principal direction or axis and $U\Sigma$ as principal components. From this interpretation, we perform $K$-means method to $m$ rows of $U\Sigma$. We suggest to apply this procedure to the estimated latent parameter tensor $\hat\Theta$. For the notational convenience, we use $\Theta$ as $\hat\Theta$ in this section. From the Tucker decomposition which is multidimensional version of SVD, we have
\begin{equation*}
\Theta = \tC\times_1\mM_1\times_2\cdots\times_K\mM_K,
\end{equation*}
where $\tC\in\mathbb{R}^{r_1\times \cdots \times r_K}$ is a core tensor, $\mM_k\in\mathbb{R}^{d_k\times r_k}$ are factor matrices with orthogonal columns, and $\times_k$ denotes the tensor-by-matrix multiplication~\cite{kolda2009tensor}. We perform mode-$k$ matricization of the tensor $\Theta$ to cluster data from mode-$k$ variable as follows.
\begin{equation*}
\Theta_{(k)} = \mM_k\tC_{(k)}\big(\mM_K\otimes\cdots\otimes\mM_1\big)
\end{equation*}
We interpret $\big(\mM_K\otimes\cdots\otimes\mM_1\big)$ as principal axes and $\mM_k\tC_{(k)}$ as principal components. By putting other mode variables as principal axes, we can exclude effects from those variables and can see $k$-th mode effect from checking principal components, $\mM_k\tC_{(k)}$. Therefore, we perform $K$-means method to $d_k$ rows of the principal components like in the matrices case.

\subsection{Clustering result on HCP}
We perform $K$-means method to cluster brain nodes based on the suggested above method. We have the estimated latent parameter $\hat\Theta$ with the rank $(23,23,8)$ which minimizes the BIC value. The number of clusters is chosen to be eight from the elbow method in figure ~\ref{figure:elbow}.
The table in \ref{table:clustering} shows eight cluster groups and their entries. Most of brain nodes fall into the cluster $\#1$ and the cluster $\#2$ which can be represented as the left side brain and the right side brain. From the group $\#3$ to the group $\#8$ , entries of each cluster share the same name encoded in the data and play a similar role in the brain. For example, the group $\#3$ and the group $\#4$ are the left and the right side of the regions called Supramarginal gyrus. Those regions are known to be in charge of interpreting tactile sensory data and involved in perception of space and limbs location \cite{carlson2012physiology,reed1994nature}. The figure \ref{figure:brain image} shows the nodes in the dataset on the brain image. Each node is connected by edges within the same cluster and the groups can be paired as the right part and the left part of the brain. The results show that the clustering based on our model successfully group the brain nodes without knowledge of the brain region.
\begin{figure}[H]
  \label{figure:elbow}
  \begin{center}
    \includegraphics[width=0.48\textwidth]{elbowmethod.png}
  \end{center}
  \caption{The elbow plot for the number of clusters}
\end{figure}

\begin{table}[H]
  \label{table:clustering}
  \center
  \begin{small}
    \footnotesize
  \begin{sc}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
Group       & \multicolumn{3}{c|}{\#1}                                                  & \multicolumn{3}{c|}{\#2}                                      \\ \hline
Brain nodes & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_L","Fpole\_L","Insula\_L",\\ "SupF\_L","CaudMF\_L",\\ "parsTriangularis\_L","parsOpercularis\_L",\\ "preCentral\_L","Tpole\_L","SupT\_L",\\ "SupT\_L","postCentral\_L", "SupP\_L",\\ "IP\_L","LO\_L","MOF\_L","SupF\_L",\\ "IsthmusC\_L","Precuneus\_L",\\ "Cuneus\_L","paraHippo\_L",\\ "Lingual\_L","SupT\_L","LO\_L"\end{tabular}} & \multicolumn{3}{l|}{\begin{tabular}[c]{@{}l@{}}"RMF\_R","Fpole\_R","Insula\_R",\\ "SupF\_R", "CaudMF\_R",\\ "parsTriangularis\_R" ,"parsOpercularis\_R",\\ "preCentral\_R", "Tpole\_R","SupT\_R",\\ "SupT\_R","postCentral\_R","SupP\_R",\\ "IP\_R","LO\_R","MOF\_R", "SupF\_R",\\ "IsthmusC\_R","Precuneus\_R",\\ "Cuneus\_R","paraHippo\_R",\\ "Lingual\_R","SupT\_R","LO\_R"\end{tabular}} \\ \hline
Group       & \#3    & \#4    & \#5   & \#6  & \#7 & \#8
\\ \hline
Brain nodes & \begin{tabular}[c]{@{}c@{}}"SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L" \\ "SupraM\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R" \\ "SupraM\_R"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_L" \\ "IT\_L" \\ "IT\_L"\end{tabular}                                      & \begin{tabular}[c]{@{}c@{}}"IT\_R" \\ "IT\_R" \\ "IT\_R"\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}"MT\_L" \\ "MT\_L" \\ "MT\_L"\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}"MT\_R" \\ "MT\_R"\\ "MT\_R"\end{tabular}                                                         \\ \hline
\end{tabular}
\end{sc}
\end{small}
\caption{Result of the brain nodes clustering: the last alphabet in each node name means the left side for `L' and the right side for `R' in the brain}
\end{table}

\begin{figure}[H]
  \centering
 \includegraphics[width = 0.45\textwidth]{brain1.jpg}
 \includegraphics[width = 0.45\textwidth]{brain2.jpg}
 \includegraphics[width = 0.45\textwidth]{brain3.jpg}
 \includegraphics[width = 0.45\textwidth]{brain4.jpg}
  \label{figure:brain image}
  \caption{Brain image of eight clusters: entries of each cluster are connected by edges}
\end{figure}

\bibliographystyle{plainnat}
\bibliography{tensor_wang}

\end{document}